Multi-modality Imagery Database for Plant Phenotyping

### 2015-04-26 datapaper_baseline.tex
Xi edit

\section{Baseline Performance}
\subsection{Our Framework}
We apply our automatic leaf segmentation, alignment, and tracking framework~\cite{yin2014a,yin2014b} to the Arabidopsis imagery to provide a baseline for future research.
Leaf segmentation aims at finding the correct number of leaves and the corresponding boundary of each leaf in each image.
Leaf alignment is to align the two tips of each leaf.
Leaf tracking is designed to track each leaf over time.

@@

As shown in Fig~\ref{fig:methodOverview}, the input of this framework is a plant video and a set of predefined templates with various shapes, scales, and orientations.
We also generate the two tips for each template for finding the corresponding tip points of each leaf via Chamfer Matching~\
First, we apply multi-leaf alignment~\cite{yin2014a} approach to find an optimal set of leaf candidates on the last frame of the video, which will provide the information of the number of leaves, tip locations and boundaries of each leaf.
Second, we apply multi-leaf tracking~\cite{yin2014b} approach, which is based on leaf template transformation, to track leaves between continuous two frames.
In the tracking process, we develop a procedure to generate new leaves and delete small leaves. 
For each frame of the video, we can generate a label image with each leaf being labeled with one color and the tip locations for each piece of leaf. 
The label for each leaf in the video maintain the same during tracking process.

\subsection{Performance Evaluation}
To evaluate the performance of leaf segmentation, alignment, and tracking, we use four criteria and provide the Matlab implementations.
Three of them are based on tip-based error, which is defined as average distance of a pair of estimated leaf tips $\hat{\bm{t}}_{1,2}$ with a pair of labeled leaf tips $ \bm{t}_{1,2}$ normalized by the labeled leaf length:

@@

We build the frame-to-frame and video-to-video correspondence respectively and generate two sets of tip-based errors.
More details can be find in~\cite{};
We define a threshold $\tau$ to operate on the corresponding tip-based errors. 
By varying $\tau$, we compute the first three criteria as:
\begin{itemize}
  \item {\it{Unmatched Leaf Rate(ULR)}}, the percentage of unmatched leaves w.r.t. the total number of labeled leaves. This can attribute to two sources. First, miss detections and false alarms. Second, matched leaves with tip-based errors larger than $\tau$.
  \item {\it{Landmark Error(LE)}}, the average tip-based errors smaller than $\tau$ of all frame-to-frame correspondent leaves.
  \item {\it{Tracking Consistency(TC)}}, the average tip-based errors smaller than $\tau$ of all video-to-video correspondent leaves. 
\end{itemize}

@@

The results of our algorithm by varying $\tau$ from 0 to 1 is shown in Fig.XXX.
And the SBD score is XXX by averaging over all plants.

>>>>

\section{Baseline Method and Performance Evaluation}
\label{sec:baseline}
\subsection{Our framework}
We apply our automatic multi-leaf alignment, segmentation, and tracking framework~\cite{yin2014a,yin2014b} to the testing set of Arabidopsis fluorescence imagery to provide a baseline for future comparison.
Leaf segmentation aims at finding the correct number of leaves and the corresponding boundary of each leaf in each image.
Leaf alignment is to align the two tips of each leaf.
Leaf tracking is designed to track each leaf over time.

@@

As shown in Fig~\ref{fig:methodOverview}, the input of this framework is a plant video and a set of predefined templates with various shapes, scales, and orientations.
To generate the template set, we first select $10$ templates with different aspect ratios from the labeled images in the training set together with the corresponding tip locations.
For each template, we scale it to $12$ different sizes in order to cover all size ranges in the database.
For each scale template, we rotated every $15^{\circ}$ to generate $24$ different directions.
Tip locations will be scaled and rotated accordingly.
Finally, we generate $2,880$ leaf templates.

%We also generate the two tips for each template for finding the corresponding tip points of each leaf via Chamfer Matching~\cite{barrow1977parametric}.
Our work is motivated by Chamfer Matching technique~\cite{barrow1977parametric}, which is used to align two edge maps.
We extend it to align multiple objects at the same time.
For each image, we use simple thresholding and edge detection to generate an edge map and mask.
First, we find the best location of each template in the edge map that has minimal chamfer matching distance, which will result in an over-complete set of leaf candidates.
Second, we apply multi-leaf alignment~\cite{yin2014a} approach to find an optimal set of leaf candidates on the last frame of the video, which will provide the information of the number of leaves, tip locations and boundaries of each leaf.
Third, we apply multi-leaf tracking~\cite{yin2014b} approach, which is based on leaf template transformation, to track leaves between continuous two frames.

In the tracking process, we develop a procedure to generate new leaves and delete small leaves.
For each frame of the video, we can generate a label image with each leaf being labeled with one color and the tip locations for each piece of leaf.
The label for each leaf in the video maintain the same during tracking process.

\subsection{Evaluation metrics}
To evaluate the performance of leaf segmentation, alignment, and tracking, we use four criteria and provide the Matlab implementations.
Three of them are based on tip-based error, which is defined as average distance of a pair of estimated leaf tips $\hat{\bm{t}}_{1,2}$ with a pair of labeled leaf tips $ \bm{t}_{1,2}$ normalized by the labeled leaf length:

@@

We build the frame-to-frame and video-to-video correspondence respectively and generate two sets of tip-based errors.
More details can be find in~\cite{};
We define a threshold $\tau$ to operate on the corresponding tip-based errors.
By varying $\tau$, we compute the first three criteria as:
\begin{itemize}
  \item {\it{Unmatched Leaf Rate(ULR)}}, the percentage of unmatched leaves w.r.t. the total number of labeled leaves. This can attribute to two sources. First, miss detections and false alarms. Second, matched leaves with tip-based errors larger than $\tau$.
  \item {\it{Landmark Error(LE)}}, the average tip-based errors smaller than $\tau$ of all frame-to-frame correspondent leaves.
  \item {\it{Tracking Consistency(TC)}}, the percentage of video-to-video correspondent leaves whos tip-based errors are smaller than $\tau$.
\end{itemize}

@@

The Matlab function for computing {\it{SBD}} is provided by~\cite{scharr2014annotated}.
The instruction on how to use the evaluation functions are included within the functions.

\subsection{Performance and analysis}
The results of our algorithm by varying $\tau$ from 0 to 1 is shown in Fig.~.
And the {\it{SBD}} score is $0.51$ by averaging over all plants.

\begin{figure}[h]
\centering
\includegraphics[width=.98\textwidth]{Figures/performance}\\
\caption{Performance of the baseline method.}
\label{fig:performance}
\end{figure}

### 2015-04-26 datapapr_labeling.tex
Xi edit

\section{Image Labeling}

Part of the dataset is manually annotated to provide ground truth tip locations, leaf segmentation results and leaf consistency overtime.
Tip locations are saved in a TXT file for each frame.
Leaf segmentation results are stored in a PNG image for each frame with one color for each leaf.
The same color is used to represent the same leaf over a sequence of frames.

We use Fluorescence images for labeling because of clear background. 
%All plants are cropped out from the background and saved as JPG files.
For Arabidopsis images, we label four frames each day.
While for bean images, we label seven frames each day because of faster leaf movement.
We develop a Matlab GUI interface for leaf labeling, as shown in Fig~\ref{fig:label}.
A user can open an image to label the two tips and annotate each leaf.
The results will be automatically saved once a user goes to label the next image.
For consistent annotation of the same leaf over time, we show a number on the center of each leaf indicating the order of labeling from the previous frame. 
%This GUI is used to annotate leaves in each frame.
%For consistent annotation of the same leaf over time, we use the labeled results of each frame to find the correspondence according to leaf centers.
%Incorrect correspondence will be manually corrected.

@@

Leaf label is implemented by clicking the boundary of one leaf at each time.
The leaf boundary is overlaid on the image for better visualization to guide the next action.
Incorrect label can be deleted right after the labeling. 
This process continues until all leaves have been annotated.
After the labeling of one plant, we visually go through the results and correct inaccurate labels. 
One example of leaf label result for one plant is shown in Fig.~\ref{fig:LabelExample}, where one color is used to represent one leaf.  
As we can see, there is leaf showing up and cover up the leaf underneath, which will not be annotated later (transition between day $4$ and day $5$).

@@

>>>>

\section{Database and Annotation}
\subsection{Statistics and protocols}
MSU-PID includes two plant databases: Arabidopsis and Bean.
The statistic information of these two databases are summarised in Tab.~\ref{tab:stat}.
The images were acquired every hour.
As there is no light at night, plants can not be captured in fluorescence and RGB color images while IR and depth cameras can still capture plants.
Therefore, in order to make it consistent, we release images captured only in day time, which are $16$ images per day for Arabidopsis and $14$ for bean for all four sensors.

\begin{table}[h]
\begin{center}
\caption{Summary of Arabidopsis and Bean databases.}
\label{tab:stat}
\begin{tabular}{c|c|c|c|c|c}
      \hline
      % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
      Database     & Subjects & Days & Images/Day & Total Images & Annotated Images \\
      \hline
      Arabidopsis &  16      &  9   &     16     &     2304     &       576    \\
      %\hline
      Bean        &   5      &  5   &     14     &     350      &       175 \\
      \hline
\end{tabular}
\end{center}
\end{table}

The two databases differs in plant resolutions.
As shown in Fig~\ref{fig:rawIm},  we grow a single bean plant while a whole tray of Arabidopsis are grown at the same time.
Therefore, the resolution of each Arabidopsis plant is much lower than that of a bean plant.
We manually crop $16$ Arabidopsis plants, which have been captured by all four sensors.
Table~\ref{tab:resolution} summaries the resolution of each plant in all four channels.

\begin{figure}[h]
\begin{centering}
\begin{tabular}{c c}
%\begin{tabular}{@{}c@{} c@{}}
%\begin{tabular}{lllllllll}
\includegraphics[width=.45\textwidth]{Figures/rawImages/a.png}&
\includegraphics[width=.45\textwidth]{Figures/rawImages/b.png}\\
  Arabidopsis & Bean \\
\end{tabular}
\caption{Color images of Arabidopsis and bean plants. }
\label{fig:rawIm}
\end{centering}
\end{figure}

\begin{table}[h]
\begin{center}
\caption{Plant resolution of Arabidopsis and Bean databases.}
\label{tab:resolution}
\begin{tabular}{c|c|c|c|c}
      \hline
      % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
      Database     & Fluorescence       & IR        & RGB      & Depth     \\
      \hline
      Arabidopsis &  ~240$\times$240 &  ~240$\times$240 & ~120$\times$120  & NA  \\
      %\hline
      Bean        & 1360$\times$1024 & 1360$\times$1024 & 1280$\times$720 & 320$\times$240    \\
      \hline
\end{tabular}
\end{center}
\end{table}

As shown in Tab.~\ref{tab:database}, MSU-PID can be used for leaf segmentation, leaf alignment, leaf tracking, and leaf counting.
To facilitate future research, we separate the database into training set and testing set.
$40\%$ of the data is used for training and $60\%$ for testing.
Specifically, $6$ plants of Arabidopisis and $2$ plants of bean are selected for training.
We will provide training and testing data in different folders.
For fair comparison, both supervised learning and unsupervised learning methods should evaluate the performance on the training and testing sets separately.
The evaluation metrics will be discussed in Sec.~\ref{sec:baseline}.

\subsection{Manual annotations}
Part of the database is manually annotated to provide ground truth tip locations, leaf segmentation results and leaf consistency overtime.
Tip locations are saved in a TXT file for each frame.
Leaf segmentation results are stored in a PNG image for each frame with one color for each leaf.
The same color is used to represent the same leaf over a sequence of frames.

We use Fluorescence images for labeling because of clear background.
For Arabidopsis images, we label $4$ frames each day.
While for bean images, we label $7$ frames each day because of faster leaf movement.
A Matlab GUI interface is developed for leaf labeling, as shown in Fig~\ref{fig:label}.
A user can open an image to label the two tips and annotate each leaf.
The results will be automatically saved once a user goes to label the next image.
For consistent annotation of the same leaf over time, we show a number on the center of each leaf indicating the order of labeling from the previous frame.
%This GUI is used to annotate leaves in each frame.
%For consistent annotation of the same leaf over time, we use the labeled results of each frame to find the correspondence according to leaf centers.
%Incorrect correspondence will be manually corrected.

@@

Leaf label is implemented by clicking the boundary of one leaf at each time.
The labeled leaf boundary is overlaid on the image for better visualization to guide the next action.
Incorrect label can be deleted right after the labeling.
This process continues until all leaves have been annotated.
After the labeling of one plant, we visually go through the results and correct inaccurate labels.
One example of leaf label result for one plant is shown in Fig.~\ref{fig:LabelExample}, where one color is used to represent one leaf.
As we can see, there is leaf showing up and cover up the leaf underneath, which will not be annotated later (transition between day $4$ and day $5$).

@@

\subsection{File types and name conventions}
We release training and testing sets in two separate folders.
In each folder, there are two subfolders named Arabidopsis and Bean.
The files in each subfolder have the following form:
\begin{itemize}
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$rgb.png: the original RGB color images;
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$fmp.png: the original fluorescence images;
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$ir.png: the original IR images;
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$depth.png: the original depth images;
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$label.png: the labeled images;
  \item plant$\_$XX$\_$day$\_$X$\_$num$\_$XX$\_$tips.txt: the labeled tip locations;
\end{itemize}
where $XX$ is an integer number indicating the specific plant information.
For each image in the database, we provide four channels images in PNG files ($\_$rgb, $\_$fmp, $\_$ir, $\_$depth).
For annotated images, we have two additional files ($\_$label, $\_$tips) saving the annotation results.
Leaf segmentation results are encoded as indexed PNG files, where each leaf is assigned a unique and consistent ID over time.
Leaf ID starts from $1$ and continuously increase.
And the background is encoded as $0$.
Tips locations are saved in TXT files where each line has the following format:
\begin{itemize}
  \item leaf ID \quad tip1$\_$x \quad tip1$\_$y \quad tip2$\_$x \quad  tip2$\_$y
\end{itemize}
where leaf ID is an integer number that is consistent with the segmentation label in the PNG file.
tip1$\_$x and tip1$\_$y are floating numbers represent the coordinates of the outer tip point.
tip2$\_$x and tip2$\_$y are floating numbers represent the coordinates of the inner tip point.

In addition to the original images and annotation results, we provide another folder named Matlab with all Matlab functions that will be used for mapping between different image channels and for performance evaluation purpose.
Note that the annotation is provided based on Fluorescence images.
In order to evaluate methods developed on other channels, we provide image mapping functions between every two channels.
